---
title: "Logistic regression"
author: "Tata iles"
date: "2023-03-10"
output: html_document
---

```{r}
library(tidymodels)
library(tidyverse)
```

# Loading data

```{r}
data <- read.csv("D:/Python-R/Databases/Telco-Customer-Churn.csv")
no_service =colnames(data)[7:14]

data <- data %>% select(-1) %>%
  mutate_if(is.character, as.factor) %>%
  mutate(SeniorCitizen=as.factor(SeniorCitizen)) %>%
  mutate_at(no_service,
            function(x) ifelse(x %in% 
                      c("No internet service","No phone service","No"),"No","Yes"))

data$Churn <- relevel(data$Churn, ref = "Yes")




#glimpse(data)
```

# Splitting the data

```{r}
set.seed(123)

data_split <- data %>% initial_split()

train <- data_split %>% training()
test <- data_split %>% testing()

```

# Model 1

```{r}
model_simple <- logistic_reg() %>% set_mode("classification") %>% 
  set_engine("glm") 

recipe <- train %>% recipe(Churn~.) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_scale(c("tenure","TotalCharges","MonthlyCharges")) 
  
  


recipe %>% prep(train) %>% juice()


wrkf1 <- workflow() %>% add_model(model_simple) %>% add_recipe(recipe)
```

## Coefficients of regression

```{r warning=FALSE}
fit(wrkf1,train) %>% tidy()

```

### Interpretation des coefficients

pour une augmentation de $exp(\beta)$ va causer une augmentation de la probabilité d'avoir "yes" par $1-exp(\beta)$

par exemple $\beta_2=-0.688$ donc une augmentation de $exp(\beta)=4.14$ cause diminution de la proba de "yes" par $1-exp(\beta)=0.497$ \<=\> 49%

## Predictions

```{r}
fitted <- last_fit(wrkf1,data_split)

collect_predictions(fitted) -> results

results
```

## Performance

```{r}

metric <- metric_set(accuracy, sens, precision,yardstick::spec,recall)


metric(results, truth=Churn, estimate=.pred_class)

fitted %>% collect_metrics()

results %>%
  conf_mat(truth = Churn , estimate = .pred_class)

results
```

```{r}
results %>%
  roc_auc(truth = Churn , estimate = .pred_Yes)
```

```{r}
results %>%
  roc_curve(truth = Churn , estimate = .pred_Yes) %>% autoplot()
```

# Model 2

## Cross validation

```{r}
data_cv <- vfold_cv(data=train, v=10 ,strata = Churn)
```

## Hyper parameter tuning using cv

```{r}
model_tune <- logistic_reg(penalty=tune(), mixture = tune()) %>%
  set_mode("classification") %>% 
  set_engine("glmnet") 

grid <- grid_regular(penalty(range = c(-2,2)),mixture(c(0,1)),levels=10)
grid

wrkf2 <- workflow() %>% 
  add_model(model_tune) %>%
  add_recipe(recipe) %>% 
  tune_grid(resamples = data_cv, grid=grid)
show_best(wrkf2,metric="accuracy")
show_best(wrkf2,metric="roc_auc")
```

## Hyper parameter tuning using bootstrap

```{r}
model_tune <- logistic_reg(penalty=tune(), mixture = tune()) %>%
  set_mode("classification") %>% 
  set_engine("glmnet") 

grid <- grid_regular(penalty(range = c(-2,2)),mixture(c(0,1)),levels=10)
grid

wrkf_bs <- workflow() %>% 
  add_model(model_tune) %>%
  add_recipe(recipe) %>% 
  tune_grid(resamples = bootstraps(train,times=10,strata=Churn), grid=grid)
show_best(wrkf_bs,metric="accuracy")
show_best(wrkf_bs,metric="roc_auc")
```

## fitting ridge regularization while tuning penalty

```{r}
model_ridge <- logistic_reg(penalty=tune(), mixture = 0) %>%
  set_mode("classification") %>% 
  set_engine("glmnet") 

grid_r <- grid_regular(penalty(range = c(0,2)),levels=10)

wrkf_r <- workflow() %>% 
  add_model(model_ridge) %>%
  add_recipe(recipe) %>% 
  tune_grid(resamples = data_cv, grid=grid_r)
show_best(wrkf_r,metric="accuracy")
show_best(wrkf_r,metric="roc_auc")
```

## fitting lasso regularization while tuning penalty

```{r}
model_lasso <- logistic_reg(penalty=tune(), mixture = 1) %>%
  set_mode("classification") %>% 
  set_engine("glmnet") 

grid_l <- grid_regular(penalty(range = c(-2,2)),levels=10)

wrkf_l <- workflow() %>% 
  add_model(model_lasso) %>%
  add_recipe(recipe) %>% 
  tune_grid(resamples = data_cv, grid=grid_l)
show_best(wrkf_l,metric="roc_auc")
```

# Model 3 (recipe change)

```{r message=FALSE, warning=FALSE}
best_model <- logistic_reg() %>% set_mode("classification") %>% 
  set_engine("glm") 

recipe1 <- train %>% recipe(Churn~.) %>% 
  step_mutate(log_avg=log(TotalCharges/MonthlyCharges))%>%
  step_BoxCox(all_numeric_predictors()) %>%
  step_other(all_nominal_predictors(),threshold =0.2) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_scale(c("tenure","TotalCharges","MonthlyCharges")) 
  
  


recipe1 %>% prep(train) %>% juice()


wrkf_best <- workflow() %>% add_model(model_simple) %>% add_recipe(recipe1)

fitted_best <- last_fit(wrkf_best,data_split)

collect_predictions(fitted_best) -> results1


```

## Performance

```{r}

metric(results1, truth=Churn, estimate=.pred_class)

results1 %>%
  conf_mat(truth = Churn , estimate = .pred_class)
```

```{r}
fitted_best %>% collect_metrics()
```

```{r}
results1 %>%
  roc_curve(truth = Churn , estimate = .pred_Yes) %>% autoplot()
```

# Comparing all models

```{r}
results1 %>% roc_auc(truth = Churn , estimate = .pred_Yes) %>% pull(.estimate) -> best
results %>% roc_auc(truth = Churn , estimate = .pred_Yes)%>% pull(.estimate) -> basic

(show_best(wrkf2,metric="roc_auc") %>% pull(mean))[1]  -> roc_tune
(show_best(wrkf2,metric="roc_auc") %>% pull(mean))[1] -> roc_bs
(show_best(wrkf_r,metric="roc_auc") %>% pull(mean))[1] -> roc_r
(show_best(wrkf_l,metric="roc_auc") %>% pull(mean))[1] -> roc_l

performance <-c(basic,roc_tune,roc_bs,roc_r,roc_l,best) %>% round(4)

data.frame(cbind(
  c("model1","model2_tune_cv",
    "model2_tune_bs","model2_r","model2_l","model3"),
  performance)) %>%
  ggplot(aes(V1,performance)) + geom_col() + coord_flip() + labs(x = "model",y="roc_auc")
```
